\thispagestyle{empty}

\section{Abstract}

This report describes the development to contribute to Ion Shell for the It project "redox"
between the winter term 2022 and summer term 2023.
First an introduction about Ion Shell is provided.
In the next chapter the typical development process which I applied, is described.
That chapter indents to showcase how documentation, testing and coding can be performed for Ion Shell.
At the end of this chapter the development process is illustrated via an examples of a feature implementation for Ion Shell.
After that chapter a reflection outlines the current status of Ion Shell.
This reflection also involves achievements of my development work and left tasks/goals I could not complete.
In the last chapter a listing of all my created issues and conducted pull requests are provided to quantify the development work.

\section{Abstract in German}

Dieser Bericht beschreibt die Entwicklung als Beitrag zu Ion Shell für das It Projekt "redox"
während dem Wintersemester 2022 und Sommersemester 2023.
Zuerst kommt eine Einleitung zu Ion Shell.
Im nächsten Kapitel wird der typische Entwicklungsprozess, den Ich angewendet habe, beschrieben.
Diese Kapitels soll aufzeigen wie Dokumentation, das Testen und die Programmierung für Ion Shell durchgeführt werden.
Zum Ende dieses Kapitels wird der Entwicklungsprozess durch ein Beispiel in Form einer Implementierung für Features verdeutlicht.
Nach diesem Kapitel soll der aktuelle Status von Ion Shell verdeutlicht werden.
Dieser Teil soll auch die Errungenschaften meiner Entwicklungsarbeit und die übrigen Aufgaben/Ziele welche Ich nicht vollendend konnte, aufzeigen.
Im letzten Kapitel wird eine Auflistung von all meinen erstellten Issues und durchgeführten Pull Requests dargestellt um
meine Entwicklungsarbeit zu quantifizieren.

\clearpage

\section{Sources of code snippets}

In several places, code snippets are used to illustrate certain points.
Some code snippets are taken from the Ion Shell repository on Gitlab \cite{ion_shell}
based on the commit id \commitid.
Unfortunately I could not generate an online link to a file based on a certain commit.
For this reason I include a local copy of the repository beside this report.
The name of this snapshot is "ion.zip".
This local copy is a snapshot of the repository based on the commit \commitid.
A source for a code snippet based on a certain file of this mentioned commit in the bibliography
is therefore a local file path within this local snapshot.

The Notation for file paths is as follows:

```
<folder> -> <folder> -> .. -> <file>
```

Therefore the following Example:

```
src -> shell -> flow.rs
```

Equals the flow.rs file inside the folder shell which is inside the folder src at the root project.

