\chapter{List of development accomplishment}

\section{Pull Requests}

\subsection{Features}
\label{accomplishment_features}

\begin{itemize}
	\item Implemented pipefail feature. If at least one command in pipeline returns an error, the whole pipeline returns an error.\cite{pr_pipefail}.
	\item Improved error message for not declared variable. Now shows the name of not found variable \cite{pr_better_not_found_variable_show}.
	\item Improved error message for invalid index. Message now shows the length of sequence where the out of bounds error occurs \cite{pr_error_message_shows_invalid_range}.
	\item Implemented work around to make file scheme work in Ion Shell inside Redox Os \cite{pr_file_scheme}
	\item Implemented subst method for Ion Shell. \cite{pr_subst_method}
\end{itemize}

\subsection{Bug fixes}\label{accomplishment_bug_fixes}

\begin{itemize}
	\item Make all buitlins of ion shell included the ion manual after building documentation \cite{pr_man_history_included_ion_manual}.
	\item Ensured that spaces between left parentheses and the first argument does not cause an error anymore .\cite{pr_fix_space_before_first_paranthese}
	\item Converted unguarded unsafe conversion to utf-8 string. \cite{pr_unsave_utf8_removal}
\end{itemize}

Related to \gls{redoxer}

\begin{itemize}
	\item Fixed semantic version of \gls{redoxer} and applied missing argument for a function to fix compiler failure \cite{pr_bump_fix_redoxer}.
\end{itemize}

\subsection{Code Quality}
\label{accomplishment_code_quality}

\begin{itemize}
	\item Scripted option to only run one selected integration test. \cite{pr_single_test_selectable}
	\item Removed redundant, more than once, termination before tokenisation. \cite{pr_removed_redundant_termination}
	\item Remove obsolete parameter in several functions in the parser section of Ion Shell. \cite{pr_remove_not_needed_parameter}
	\item Fixing warnings by using modern API of criterion in benchmarks. \cite{pr_fix_warnings_deprecated_criterion_api}
	\item Added unit tests for parsing let statements. \cite{pr_added_unit_test_let_grammar}
	\item Added unit tests for terminating lines before tokenisation. \cite{pr_test_terminator_todos}
	\item Removed not used struct field. \cite{pr_removed_flag_undead_code}
\end{itemize}

\subsection{Documentation}\label{accomblishment_docs}

\begin{itemize}
	\item Moved msrv from clippy file to Cargo.toml. \cite{pr_moved_msrv_to_cargo}
	\item Documented how the documentation of builtin are generated into the Ion Shell online manual. \cite{pr_docs_how_builtin_generated}
	\item Wrote section about how integration test work and how to write them. \cite{pr_docs_section_integraion_test}
	\item Provided additional command for building ion manual locally. \cite{pr_docs_commands_building_ion_manual}
	\item Fixed formatting in the builtin section of the ion manual \cite{pr_docs_fix_wrong_builtin_manual_section}
	\item Adjusted msrv to sync it within the project \cite{pr_docs_adjust_msrv_project}
	\item Added notes/links to ion plugin. \cite{pr_docs_ion_plugin}
	\item Documented feature of initializing file. \cite{pr_docs_init_file}
	\item Added output to example of creating an array withing the ion manual \cite{pr_docs_output_create_array}
	\item Created a section about how integration tests work in Ion Shell. It also contains guides how to write integration tests for Ion Shell \cite{pr_docs_how_to_integration_test}.
\end{itemize}

\subsection{Configuration}

\begin{itemize}
	\item Make Linux job work in CI and fixed version in MakeFile. \cite{pr_conf_partial_ci_fix}
	\item Removed always overwritten test files and made it ignored in git. \cite{pr_conf_overwritten_files}
	\item Updated section about joining the redox community on matrix server. \cite{pr_docs_update_community_links}

\end{itemize}

\section{Issues}

Resolved issues mean that they could be closed due to my contributions.

\subsection{Opened and resolved Issues}

\begin{itemize}
	\item Reported problem of not working builtin documentation generation within ion manual \cite{issue_builtin_generation}. Solved by pull request \cite{pr_man_history_included_ion_manual}.
	\item Reported bug that white spaces between opening parentheses and the first argument causes an error \cite{issue_spaces_between_opening_parentheses}. Solved by pull request \cite{issue_spaces_between_opening_parentheses}.
	\item Reported that ion manual is not updated by the documentation of the builtin methods \cite{issue_ion_manual_not_updated}. Solved by pull request \cite{ion_shell_subst_ion_subst_pr}.
	\item Reported that the test suite of Ion Shell does not compile with intended rust version \cite{issue_bug_test_not_compile}. Solved by pull request \cite{pr_conf_partial_ci_fix}.
\end{itemize}

\subsection{Resolved Issues}

The following issues were not opened by me but resolved by me via contribution.

\begin{itemize}
	\item Feature request of new builtin method subst to allow for replacing empty arrays with an default array \cite{issue_subst_request}. Resolved by pull request \cite{pr_subst_method}.

\end{itemize}

\subsection{Opened issues}

This issues were opened by me to report an problem. But they were not solved by me.

Related to \gls{redoxer}

\begin{itemize}
	\item Requested the rebuilding of the docker image for \gls{redoxer} so that the Ion Shell repository can be build with a new rust version \cite{issue_request_new_docker_version}.
\end{itemize}

\subsection{Still Opened Issues}

This issues were opened by me to report an problem. But they have not been solved yet.

\begin{itemize}
	\item Noted null-coalescing operator as a possible language feature \cite{issue_null_coalescing_operator}. Was brought up in the discussion in the issue \cite{pr_subst_method}.
	\item Noted down the trenary as a possible language feature \cite{issue_ternary_operator_arrays_strings}. Was brought up in the discussion in the issue \cite{pr_subst_method}.
	\item Reported the missing feature of handling redox os schemes in general \cite{issue_disk_scheme}.
	\item Proposed more information for error messages to end user. Espeacally by providing the locatoin where the error occured, ake line and colum \cite{issue_enrich_error_message_information}.
	\item Reported general failing of Ci for ion shell \cite{issue_failing_ci}.
	\item Reported that a job called "linux:stable" in the Ci still fails \cite{issue_ci_linux_stable_fail}.
\end{itemize}

Related to \gls{redoxer}.

\begin{itemize}
	\item Reported that command "redoxer exec true" does fail on fedora \cite{issue_redoxer_exec_true}.
\end{itemize}


